\documentclass{article}
\usepackage{tikz}
\usepackage{bm}
\usepackage{listings}
\usepackage{amsfonts}
\usepackage{amsmath}

\lstset{
  language=Python,
  basicstyle=\ttfamily,
  keywordstyle=\color{blue},
  commentstyle=\color{green!40!black},
  stringstyle=\color{red},
  numbers=left,
  numberstyle=\tiny,
  stepnumber=1,
  numbersep=5pt,
  backgroundcolor=\color{gray!10},
  frame=single,
  rulecolor=\color{black!30},
  tabsize=2,
  captionpos=b,
  breaklines=true,
  breakatwhitespace=false
}
\begin{document}

\newcommand{\lettercircle}[1]{%
  \begin{tikzpicture}
    \node[circle, draw, minimum size=1cm] at (0,0) {#1};
  \end{tikzpicture}%
}


\section*{Problem Description: Counting Independent Sets in a Tree}
In this algorithm project, our primary objective is to develop an efficient algorithm that works in linear time to count independent sets within a tree structure. Independent sets, in the context of graph theory, refer to subsets of vertices in which no two vertices are adjacent. In our case, we focus on trees, which are a special type of graph with no cycles.

\subsection*{Solution Description:}
We utilized the depth fisrt search method. Traverse the nodes from the leaves to the root(bottom-up).
For each node, consider a subtree, whose root is that node, to get the IS of that subtree, denoted as $S$, we divide it two subsets $U$ and $C$: 
\begin{enumerate}
  \item IS without picking that node. denote as $U$ (unchosen), the count number of it, it's the cardinality of the set, denoted as $|U|$, the recursive procedure to get it is combinationsWithoutCurrent,  
  \item IS with the node, denote as $C$ (chosen), the count of it is $|C|$, the procedure is combinationWithCurrent . 
  \item the all IS,  denoted as $S$ , count number of all IS is $|S|$, the procedure is getAllSetsCount.
  Intuitively, the count number of the IS has formula as below:
    \[ |S| = |C| + |U| \] 
\end{enumerate}
We will calculate above three states for each node recursively, until we reach the root, the number of all IS count of the root,
$|S_{root}|$, is exactly the output expected. Pseudo code in python is as below: 
\newpage
\begin{lstlisting}
def combinationsWithoutCurrent(current):
    num = 1
    for child in current:
        num *= child.cachedWithoutValue + child.cachedWithValue
    return num


def combinationsWithCurrent(current):
    num = 1
    for child in current:
        num *= child.cachedWithoutValue
    return num


def getAllSetsCount(arrayOfNodes):
    for node in arrayOfNodes:
        with_current = combinationsWithCurrent(node)
        without_current = combinationsWithoutCurrent(node)
        node.cachedWithValue = with_current
        node.cachedWithoutValue = without_current
    # return root node calculated value [-1] is the last element which is root in this case
    return arrayOfNodes[-1].cachedWithValue + arrayOfNodes[-1].cachedWithoutValue
\end{lstlisting}

The tree and its node definitions are as below:

\begin{lstlisting}
class TreeNode:
    def __init__(self, data):
        self.data = data
        self.cachedWithValue = 0
        self.cachedWithoutValue = 0
        self.children = []

class Tree:
    def __init__(self, root_data):
        self.root = TreeNode(root_data)
\end{lstlisting}

Tree is nothing but a root tree node where we can end and all IS count of the tree is exactly $S_{root}$.

\subsection*{Proof of correctness and time complexity analysis}
We traverse the tree node exactly once, we denote the number of all nodes in the tree as N. So the time complexity of the algorithm with Big O notation is $O(N)$

Proof is as in the below, we define some notations: recall three states from the above, we denote sets $U_{L}$, $C_{L}$, $S_{L}$, as each set for the node L, we consider the bottom up recursion, using the induction method:
\begin{itemize}
  \item Base case: Consider the case, a leaf node, denoted as $L$, a leaf is a node who has no children, the IS of $L$ without picking it, is $\emptyset$, $|U_{L}=1|$  the IS of $L$ including it, is the node itself${L}$, $C_{L}=1$, all IS set count is: $|S_{L}| = |U_{L}| + |C_{L}| = 2$ 
  \item Inductive Hypothesis: Because we traverse from the bottom to the top, let's move one level up, consider the subtree starting from the parent of the node, let's denote the parent node as $P$. To calculate all IS Count of the subtree roots from $P$, let's divide into two cases: 
    \begin{enumerate}
      \item IS without P
      \item IS with P. 
    \end{enumerate}

    In case of 1, think about each child node, denoted as $D_{1...n}$, meaning Nth child node of the parent node P, from 1st to Nth,
    let's consider all Independent sets of each substree roots from the child node, 
    \[ S_{D} =  U_{D} \cup C_{D} \tag{$*$} \] 
    Because there is no P in the any set of $S_{P}$, recall the definition of independent set, we have no limit in combine all the sets inside each child IS set.
    We consider Cartesion product of each $S_{D_n}$, we denote $a_{D_n}$ as any element in $S_{D_n}$:
    \[ U_P =  S_{D_1} \times S_{D_2} \times \ldots \times S_{D_n} = \{(a_{D_1}, a_{D_2}, \ldots, a_{D_n}) \mid a_{D_1} \in S_{D_1}, a_{D_2} \in S_{D_2}, \ldots, a_{D_n} \in S_{D_n}\} \]
    Here, the each pair does not need be ordered.

    The cardinality of the Cartesian product $S_{D_1} \times S_{D_2} \times \ldots \times S_{D_n}$ is the product of the cardinalities of the individual sets:
    \[|U_P| =|S_{D_1} \times S_{D_2} \times \ldots \times S_{D_n}| = |S_{D_1}| \cdot |S_{D_2}| \cdot \ldots \cdot |S_{D_n}| \]

    In case of 2, due to the fact the parent node is included in each set, similarily, 
    the result will be the combinations of IS of each child without the child node inside (set of $U$), hence:
    \[ C_P =  U_{D_1} \times U_{D_2} \times \ldots \times U_{D_n} = \{(a_{D_1}, a_{D_2}, \ldots, a_{D_n}) \mid a_{D_1} \in U_{D_1}, a_{D_2} \in U_{D_2}, \ldots, a_{D_n} \in U_{D_n}\} \]
    \[|C_P| =|U_{D_1} \times U_{D_2} \times \ldots \times U_{D_n}| = |U_{D_1}| \cdot |U_{D_2}| \cdot \ldots \cdot |U_{D_n}| \]

    recall equation $*$, in the case one, to get $S_{D}$, we already get $U_{D}$ first, due to the cache, we don't need compute it again. 

    hence, the case two does not add extra computaltional complexity.

    Above two cases are exactly what described in the pseudo code.
    
    The final step:
    \[ S_{P} =  U_{P} \cup C_{P} \] 
    \[ |S_{P}| =  |U_{P}| + |C_{P}| \] 



  \item Inductive Step: For each level up, we do the exactly as Inductive Hypothesis, until reach the root. $|S_{root}|$, is exactly the output expected.

  \end{itemize}


\subsection*{Input and output}
Input: a text file, 
A nested list to represent a tree:
for example:
\[ [1 [2 [4, 5], 3 [6, 7]] ] \]
representing a tree as below:


\begin{tikzpicture}
  \node {1}
    child[sibling distance=3cm]{
      node {2}
      child [sibling distance=1cm]{
        node {4}
      }
      child [sibling distance=1cm]{
        node {5}
      }
    }
    child [sibling distance=3cm]{
      node {3}
      child [sibling distance=1cm]{
        node {6}
      }
      child [sibling distance=1cm]{
        node {7}
      }
    };
\end{tikzpicture}

where each bracket represents a level in the tree. the nodes in the same level (the same bracket) are siblings.
the first node in the fisrt level (first bracket), represents the node of the tree.

The expected output of this tree is 41, a number represents the count of the independent sets in the tree. 


Another example:

\[ [10 [20 [40, 50], 30 [60 [80, 90], 70]]] \]
In this representation:

10 is the root.
20 and 30 are its children.
40 and 50 are children of 20.
60 and 70 are children of 30.
80 and 90 are children of 60.
This is a simple way to represent a tree structure using nested lists. Each sublist represents a node in the tree, where the first element is the node value, and the subsequent elements (if any) are the children of that node. 

In the input file, there is a nested list in each line, each list represents a tree, such as:
\[ [1 [2 [4, 5], 3 [6, 7]]]\]
\[[50 [30 [20, 40], 70 [60, 80]]]\]
\[[8 [3 [1, 6 [4, 7]], 10 [9, 14]]]\]
\[[1 [2 [4 [8, 9], 5 [10, 11]], 3 [6 [12, 13], 7 [14, 15]]]]\]

In the output file, each line representing a number, the result of number of IS in each correspondinh tree from the input file:
\[10\]
\[14\]
\[26\]
\[42\]
Input file will be input.txt under the current directory while output file will be output.txt.



\end{document}



